---
title: "Train and evaluate models with tidymodels"
date: "`r Sys.Date()`"
output:
  officedown::rdocx_document:
    reference_docx: C:/Users/George Oduor/Documents/template.docx   
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
        style: Table Caption
        pre: 'Table '
        sep: ': '
      conditional:
        first_row: true
        first_column: false
        last_row: false
        last_column: false
        no_hband: false
        no_vband: true
    plots:
      style: Normal
      align: center
      caption:
        style: Image Caption
        pre: 'Figure '
        sep: ': '
  word_document:
    toc: yes
  html_document:
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    theme: united
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width = 8, fig.height = 5)
```

# Introduction

The goal of this analysis is to **predict with precision the probability of customer churn** and consequently *help in customer retention and mitigation of the risks related to customer churn.*

## Business Value

Below are some the business values that this analysis will create:

1\. Reduce churn rate by detecting early clients who are about to churn.

2\. Increase customer retention.

3\. Understand customer base better based on their activity.

## Data-set

The data set used in this project was sourced from client transaction data recorded since the inception of the banking institution.Based in this data set,informative feature sets were created which were subsequently used in the predictive model.

### Inclusion Criteria

The clients records included in this study were selected from those who were determined to have done at least one transaction within the banking platform.

### Exclusion Criteria

Clients who on boarded the digital platform of the bank and did not make any transaction.

## Project setup and data loading

```{r introduction cell,warning=F,message=F,error=F}
library(tidymodels);library(feather);library(tidyverse);library(janitor);library(flextable);library(SmartEDA);library(corrplot)
dataset <-read.csv("E:/New folder/denis/old/data/rawdata/churn.csv") %>% 
  # mutate_at(c("Min_LoanTaken","Max_LoanTaken","FirstTrxDate","ProductID"),as.numeric) %>% 
  select(-c(X)) 

```

# Exploratory Data Analysis

## Data Overview

```{r}
ExpData(dataset) %>% 
  flextable() %>% 
  autofit() 
```

**Observations and inferences.**

-   The data set contains `r format(nrow(dataset),big.mark=",")` records of customer transactions with `r length(names(dataset))` features.

-   98.51% of the cases have complete records meaning the remaining 1.49 % have missing values.This is normal with most real life data sets that contain missing values.

## Uni-variate Data Analysis.

### Target Variable

```{r target var}

dataset %>% 
  group_by(Churn) %>% 
  tally(name = "Count") %>% 
  mutate(Percent = paste(round(Count/sum(Count),2)*100,"%")) %>% 
  adorn_totals('row') %>% 
  flextable() %>% 
  autofit() 
```

**Observations and Inferences**

-   The institution has suffered a 22% churn rate overtime.

### Explanatory Variables

#### Categorical Variables

```{r}
dataset %>% select(-Churn) %>% ExpCatViz(theme = theme_classic())
```


```{r}
ExpCTable(dataset %>% select(-Churn) %>% 
            mutate(HasFixedDeposit = ifelse(HasFixedDeposit ==1,"Yes","No"))) %>% 
  #pull(Variable) %>% unique() %>% paste(collapse = ",")
  flextable() %>% autofit() %>% merge_v(j=1,target = "Variable")
```

__Observation & inferences__

- _Number of products_ : At least half (54%) subscribed to at least two of the financial institution's digital products while a very small percentage subscribed to more than 2 products.

- _HasFixedDeposit_ : Minority of the clients,1.72%, have subscribred to fixed deposit product of the institution.



\newpage

<!---BLOCK_LANDSCAPE_START--->
#### Numeric Variables

```{r}
num_vars <- dataset %>% select(-c(Churn))
num_vars %>% 
  # select(Tenure,No_of_transactions) %>% 
  ExpNumStat() %>% select(-c(Group,NegInf,PosInf,Per_of_Missing,CV,nNeg,nZero,NA_Value,nPos)) %>% #view()
  flextable(theme_fun = set_flextable_defaults(font.size = 12,padding = 0)) %>% autofit()

```

```{r}
# dataset %>% select(is.numeric)
```

__Observations and inferences__

. 
<!---BLOCK_LANDSCAPE_STOP--->

## Bivariate Distribution

### Categorical Variables Vs Target Variables

```{r}
dataset %>% 
  mutate(HasFixedDeposit= ifelse(HasFixedDeposit == 1,"Yes","No")) %>% 
  select(HasFixedDeposit,ProductsSub,Churn) %>% 
  reshape2::melt('Churn') %>% 
  group_by(Churn,variable,value) %>% 
  tally() %>% 
  reshape2::dcast(variable+value~Churn,value.var = "n") %>% 
  group_by(variable) %>%
  split(.$variable) %>% 
  map_df(~adorn_totals(.,c("row","col")) %>% 
           mutate_all(.funs = function(x)ifelse(is.na(x),0,x)) %>% 
           mutate(ChurnRate = paste(round(Yes/Total,2)*100,"%")) ) %>% 
  mutate_all(format,big.mark=",") %>% 
  flextable(defaults = set_flextable_defaults(font.size = 8)) %>% 
  autofit() %>% 
  merge_v(j = 1,target = 'variable') %>% 
  align_text_col(align = "right")
  
  
```

__Observations and inferences__

- All clients (100%) subscribed to only one product appeared to have churned as compared to clients subscribed to at least 2 or more products.
- It is less likely for clients who are presently servicing theor loans to churn as compared to other clients.This might be caused by the fact that the institution's primary product is loans oriented.



<!---BLOCK_LANDSCAPE_START--->
### Numerical Variables vs target variable

```{r}
options(scipen = 999)
num_vars <- dataset %>% select(names(num_vars),Churn)
ExpNumStat(num_vars,by = "G",gp = 'Churn') %>% 
  select(-c(NegInf,PosInf,Per_of_Missing,CV,nNeg,nZero,NA_Value,nPos)) %>% 
  flextable() %>% autofit()
```


<!---BLOCK_LANDSCAPE_STOP--->

### Identical feature removal 

### Explanatory Variable Correlations

```{r}
corr_stats <- dataset %>%
  select_if(is.numeric) %>%#
  select(contains('Loan')) %>% 
  cor()

corrplot(corr_stats, type = "lower",
         tl.col = "black", tl.srt = 45)  

```

## Base Model

```{r}
tidymodels_prefer()
set.seed(1)
churn_sc_data <- initial_split(dataset,prop = .8,strata = "Churn")

train <- training(churn_sc_data)
test <- testing(churn_sc_data)

library(caret)

```

## Feature engineering

```{r}
lm <- logistic_reg() %>% 
  fit(Churn~.,data=dataset %>% mutate(Churn=as.factor(Churn)))

```

