

# Dataset Overview

```{r dataimport and library ,message=F,warning=F,error=F}
# libraries
library(tidyverse);library(janitor);library(flextable)
library(dlookr);library(feather);library(DataExplorer)
# dataset
datafile <- read_feather("data/input/customers.feather") %>% 
  select(-New_Status,-AMount_Mobilised,-ProductID)
```
## Data Quality

### Dataset Dimensions.
```{r}
plot_intro(datafile,ggtheme = theme_bw(),geom_label_args = list(size=4))
```


```{r}
introduce(datafile) %>% 
  t() %>% 
  as_tibble(rownames = "Variable") %>% 
  rename('Values'="V1") %>% 
  filter(Variable != "memory_usage")
```

```{r dimensions}
df_dim = dim(datafile)
```

The dataset under study had `r df_dim[1]` records and `r df_dim[2]` variables inclusive of target variable.

### Variable Explanations.

```{r }
tibble(
  ClientID = "Unique Client Identifier",
  CreatedOn = "Date when the Account Was created",
  No_ofProducts = "Number of bank products the client is subscribed to",
  Tenure = "The length of time a client has stayed with the institution since onboarding in days",
  GenderID = "Clients Gender",
  FirstTrxDate = "The first time the client made a transaction after onboarding",
  LastTrxDate = "The last time the client made a transaction after onboarding",
  TrxTenure = "Length of time in days between the first and the last transaction",
  Lag_Debit = "Difference between the current withdrawal and the preceeding withdrawal",
  Lag_Credit = "Difference between the current deposit and the preceeding deposit",
  No_Trx = "Number of transactions the client had done with the bank",
  WeekendTrx = "Number of transactions done on weekends",
  EndMonthTrx = "Number of transactions done at the end of the month",
  MidMonthTrx = "Number of transactions done at the mod of the month",
  TransactedProducts = "Number of products the client has actualy transacted after subscription",
  Transactions_per_week = "Number of transactions a client made on a weeekly basis",
  Transactions_per_month = "Number of transactions a client made on a monthly basis",
  Transactions_per_year = "Number of transactions a client made on a yearly basis",
  Transactions_per_day = "Number of transactions a client made on a daily basis",
  DepositTrx = "Number of deposit transactions",
  DepositTrx_per_day = "Number of deposit transactions made on a daily basis",
  DepositTrx_per_week = "Number of deposit transactions made on a weekly basis",
  DepositTrx_per_month = "Number of deposit transactions made on a monthly basis",
  DepositTrx_per_year = "Number of deposit transactions made on a yearly basis",
  Total_Deposits = "Sum of deposit amount made",
  Total_Deposits_per_day = "Sum of deposit amount saved on a daily basis",
  Total_Deposits_per_week = "Sum of deposit amount saved on a weekly basis",
  Total_Deposits_per_month = "Sum of deposit amount saved on a monthly basis",
  Total_Deposits_per_year = "Sum of deposit amount saved on a yearly basis"
) %>% 
  t() %>% 
  as_tibble(rownames = "Variable") %>% rename("Explanation"="V1")
```


### Quality

```{r}
dlookr::diagnose(datafile) %>%
  as.data.frame() %>% arrange(desc(missing_count)) %>% 
  mutate_at(.vars=names(select(.,-variables,-types)),.funs = function(x)round(x,2))
```
 __Observations and Inferences__
 
From the above ,it was noticed that a number of variables had missing values.The missing values occurrence where however systematic showing an indication that the client had never used the service/product before.These will later require imputation.



# Exploratory Data Analysis.

The main interest under exploratory analysis is to understand the underlying distrubutions on every variable and their interactions.
## Univariate Analysis.
### Target Variable

```{r}
datafile %>% tabyl(ChurnStatus) %>% adorn_totals("row") %>% as_tibble() %>% 
  mutate(percent = round(percent,2))
```

```{r}
plotcol = "purple"
plots_theme <- theme(text = element_text(size=12),
                     plot.title = element_text(size = 12,hjust = .5,face = "bold"))
datafile %>% tabyl(ChurnStatus) %>% 
  ggplot(aes(ChurnStatus,percent)) +
  geom_col(aes(fill = factor(c("green","red"),labels = c("Churn","Not Churn"))))+
  plots_theme+theme(legend.position = 'none')+
  labs(title = "Churn Status")
```
__Observations and inferences__

As we can observe from above,the institution had a 35% churn rate.

### Explanatory Variables
#### Categorical Variables

```{r}
discretevars = datafile %>% select(-ChurnStatus) %>% 
  summarise_all(.funs = function(x)length(unique(x))) %>% 
  t() %>% 
  as_tibble(rownames = "Variable") %>% 
  filter(V1<=11) %>% pull(Variable)

datafile  %>% select(discretevars) %>% 
  mutate_all(as.factor) %>%
rowid_to_column() %>% 
  reshape2::melt(id.vars="rowid") %>% 
 group_by(variable,value) %>% 
  tally() %>% 
  ggplot(aes(value,n))+
  geom_col(fill=plotcol)+
  facet_wrap(.~variable,scales = 'free',nrow = 3)+
  plots_theme
```
```{r}
# datafile  %>% 
#   select(ChurnStatus,discretevars) %>% 
#   mutate_all(as.factor) %>%
#   rowid_to_column() %>% 
#   reshape2::melt(id.vars=c("rowid","ChurnStatus")) %>% 
#   group_by(ChurnStatus,variable,value) %>% tally() %>% 
#   reshape2::dcast(variable+value~ChurnStatus,value.var = 'n',
#                   fun.aggregate = sum,na.rm = T) %>% 
#   split(.$variable) %>% 
#   map_df(~adorn_totals(.,'row')) 
```

__Observations and inference__

1. Majority of customers have subscribed for 2 products.
2. Most customers spear to be males interns of gender
3. Of the registered customers,majority have transacted in more than 2 products.
4. Majority of the customers have no fixed deposit with the bank.
5. Most clients who have transacted dint have  a running loan at the time pf the analysis
6. A bigger percentage of customers have never taken loans with the bank.

#### Numerical Variables

```{r}
numvars <- datafile %>%
  select(-discretevars,-CreatedOn,-ClientID,-contains('Date'))

numvars %>% select(-contains(c('loan','HasFixedDeposit','Total_Withdrawals',
                               'Max_Withdrawals',
            'Av_Withdrawals','days_since_last_trx','LastDepositBalance',
            'WeekdayTrx','EndMonthTrxs','Total_Deposits_per_month',
            'Total_Deposits_per_year','Transactions_per_month',
            'Transactions_per_year','TransactedProducts'))) %>% 
  reshape2::melt(id.vars = "ChurnStatus") %>% 
  group_by(ChurnStatus,variable) %>% 
  summarise(
    sum = round(sum(value,na.rm = T),2),
    mean = round(mean(value,na.rm = T),2),
    variance = round(var(value,na.rm = T),2),
    sd = round(sd(value,na.rm = T),2),
    min = round(min(value,na.rm = T),2),
    max = round(max(value,na.rm = T),2),
  ) %>% 
  arrange(variable) %>% 
  select(variable,everything()) %>% 
  kableExtra::kable() %>% kableExtra::kable_styling()
  # reshape2::melt(id.vars = c('ChurnStatus','variable')) %>% 
  # clean_names("upper_camel") %>% 
  # reshape2::dcast(Variable+Variable_2~ChurnStatus,value.var = 'Value',fun.aggregate = sum) %>% 
  # view()
```

##### Deposits transactions count Related

```{r}
datafile %>%
  select(contains('deposittrx')) %>% 
  rowid_to_column() %>% 
  reshape2::melt('rowid') %>% 
  ggplot(aes(log(value+1)))+
  geom_histogram(fill=plotcol)+facet_wrap(.~variable,scales = 'free')+
  plots_theme
```


##### Deposit transaction amount Related 

```{r}
numvars %>% 
  select(contains('Total_Deposits'),
         Max_Deposits,Av_Deposits,Var_Deposits,sd_Deposits) %>% 
  rowid_to_column() %>% 
  reshape2::melt('rowid') %>% 
  ggplot(aes(log(value+1)))+
  geom_histogram(fill=plotcol)+
  facet_wrap(.~variable,scales = 'free',ncol = 3,nrow=4)

# datafile
```

##### Transaction Count Related

###### Descriptive Statistics

```{r}
numvars %>% 
  select(No_Trx,WeekendTrx,WeekdayTrx,EndMonthTrxs,MidMonthTrxs,
         contains('Transactions_per')) %>% 
  describe()
```


###### Visualization

```{r}
numvars %>% select(No_Trx,WeekendTrx,WeekdayTrx,EndMonthTrxs,MidMonthTrxs,contains('Transactions_per')) %>% 
    rowid_to_column() %>% 
  reshape2::melt('rowid') %>% 
  ggplot(aes(log(value+1)))+
  geom_histogram(fill=plotcol)+facet_wrap(.~variable,scales = 'free')
```
__Observations and Inferences__

- Number of transactions appear to be right skewed at Mean of 22.41 and sd = 41, transactions.Most clients have done 1-3 transactions only.

- Yearly transactions were observed to be evenly distributed across the year.

- Most customers make at most 3 transactions per month and less than 2 per week.

##### Withdrawals Related

###### Descriptive Statistics table

```{r}
numvars %>% select(Total_Withdrawals,Max_Withdrawals,Av_Withdrawals,Var_Withdrawals,sd_Withdrawals) %>% describe() %>% 
  mutate_at(.vars = names(select(.,-variable)),.funs = function(x)round(x,2))
```

###### Visualization

```{r}
numvars %>% select(Total_Withdrawals,Max_Withdrawals,Av_Withdrawals,Var_Withdrawals,sd_Withdrawals) %>% 
    rowid_to_column() %>% 
  reshape2::melt('rowid') %>% 
  ggplot(aes(log(value+1)))+
  geom_histogram(fill=plotcol)+facet_wrap(.~variable,scales = 'free')
```
## Bivariate Analaysis
### Target Variable Vs Categorical Variables

#### Descriptive Statistics

```{r}
target_cat_sum <- datafile %>% 
  select(discretevars,ChurnStatus) %>% 
  rowid_to_column() %>%
  reshape2::melt(id.vars=c('rowid',"ChurnStatus")) %>% 
  group_by(ChurnStatus,variable,value) %>% 
  tally() %>% 
  reshape2::dcast(variable+value~ChurnStatus,value.var = "n",
                  fun.aggregate = sum,na.rm=T) %>% 
  split(.$variable) %>% 
  map_df(~adorn_totals(.,c('row','col')) %>% 
           mutate(ChurnRate = paste(round(Churn/Total * 100,2) ,"%"))) %>%
  as_tibble() 
target_cat_sum
```


```{r}

plots <- target_cat_sum %>% 
  mutate(ChurnRate = parse_number(ChurnRate)) %>% 
  split(.$variable) %>% 
  map(~ggplot(.,aes(variable,ChurnRate,fill = value))+
  geom_col(position=position_dodge(width=1))+
  # labs(title = paste("Churn Rate Vs ",variable))+
  plots_theme)

for (i in plots) {
  print(i)
}
```

### Target Variable Vs Numerical Variables

##### Deposits transactions count Related

###### Descriptive Statistics

```{r}
datafile %>% 
  select(contains('deposittrx'),ChurnStatus) %>% 
  group_by(ChurnStatus) %>% 
  describe() %>% 
  mutate_at(.vars = names(select(.,-variable,-ChurnStatus)),
            .funs = function(x)round(x,2))
```


###### Visualization

```{r}
datafile %>% 
  select(contains('deposittrx'),ChurnStatus) %>% 
  rowid_to_column() %>% 
  reshape2::melt(c('rowid','ChurnStatus')) %>% 
  ggplot(aes(log(value+1),fill=ChurnStatus))+
  geom_density()+facet_wrap(.~variable,scales = 'free')
```
### Correlation Analysis.

In order to plot correlation,features that exhibited some distributions were selected.

```{r}
selected_num_vars <- datafile %>% 
  select(Total_Deposits,Total_Deposits_per_day,Total_Deposits_per_week,
         Total_Deposits_per_month,Total_Deposits_per_year,Av_Deposits,
         No_Trx,WeekdayTrx,WeekdayTrx,EndMonthTrxs,MidMonthTrxs,
         Transactions_per_day,Transactions_per_week,Transactions_per_month,
         Transactions_per_year,Total_Withdrawals,Max_Withdrawals,Av_Withdrawals,
         days_since_last_trx,LastDepositBalance)
correlation <- cor(selected_num_vars)
corrplot::corrplot(correlation,method = "pie",type = "lower",title = "Correlation Matrix",diag = T)
```
Observations and Inferences

- Some variables are highly correlated with others. e.g. Total withdrawals vs Total deposits.
- This correlation will be dealt with at feature engineering phase.

__Variable Selection for the next stage of analysis__

```{r}
modelData <- 
datafile %>% select(discretevars,ChurnStatus) %>% 
  mutate(
    GenderID = ifelse(is.na(GenderID),'Unknown',GenderID),
    HasActiveLoans = ifelse(is.na(HasActiveLoans),"NO",HasActiveLoans)) %>% 
  mutate_all(as.factor) %>% 
  bind_cols(selected_num_vars)

write_feather(modelData,"data/output/model_data.feather")
```



